---
title: "Final_data_extractor"
author: "oscar"
date: "2025-06-18"
output: html_document
---

In this script we are going to extract the final threshold and contrast for each durations of each tasks for each subjects of the discrimination experiment. we will then merged these data into a big dataframe that would be useable for statistical analysis.


# Importing library
 we import here every library that we want to use
```{r}
library(ggplot2)
library(tidyverse)
library(readr)
library(purrr)
library(dplyr)
# we should add some for the stats but we are not here yet
```
# Importing the data of each participants


```{r}
base_path <- "/home/ruiz/Documents/Stage_brxl/DATA_BOCH_LAW/DIS"
tasks <- c("landscapes", "faces", "objects")
participants <- c("01", "02", "03", "04","05","06", "07" , "08" ,"09","10","12","13","14" ,"15", "16", "17")

for (version in participants) {
  for (category in tasks) {
    # Construct file path
    file_path <- file.path(base_path, 
                          paste0("subj_",version, "_bloch_law_", category, "_discr"),
                          paste0("BL_", category, "_", version, "_data_table.csv"))
    
    # Create variable name
    var_name <- paste0(toupper(substr(category, 1, 1)), "_", version)
    
    # Import and assign
    assign(var_name, read_csv(file_path))
    message("Imported: ", var_name)
  }
}

```
Add a "task" column that contain F, L or O depending of the task (faces, landscapes, objects)

```{r}

add_task_column <- function(df, category) {
  df %>% mutate(task = category)
}

prefix_task_map <- list(
  "L" = "landscapes",
  "F" = "faces",
  "O" = "objects"
)


all_vars <- ls(pattern = "^(L|F|O)_[0-9]{2}$")


for (var_name in all_vars) {
  prefix <- substr(var_name, 1, 1)  
  task <- prefix_task_map[[prefix]]  
  assign(var_name, add_task_column(get(var_name), task))
}

```



## trials 500 to 50 per durations

```{r}
#A function to split the data in groups of durations
split_by_factor <- function(data, factor_var) {
  # Convert variable to factor if it isn't already
  if (!is.factor(data[[factor_var]])) {
    data[[factor_var]] <- factor(data[[factor_var]])
  }
  
  # Split data frame into list of subsets
  split(data, data[[factor_var]])
}


#we are calling every object that got a name with L,F,O then _ and a number so we are calling the name of each of our dataset
dataset_names <- ls(pattern = "^[LFO]_[0-9]{2}$")

#we create an empty list so we can fill it with the result of process_split_trials
split_results <- list()

# we make a loop to split each dataset in durations groups
for (ds_name in dataset_names) {
  dataset <- get(ds_name)
  
  split_data <- split_by_factor(dataset, "duration")

  result_name <- paste0(ds_name, "_split_by_", "duration")
  split_results[[result_name]] <- split_data
}



# weird deepseek function (use a lot of tidyverse and dplyr and I dont know much about it) but it seem to work will have to look into it more indeepth 
process_split_trials_safe <- function(split_data) {
  # Recursive function to handle nested lists
  modify_structure <- function(x) {
    if (is.data.frame(x)) {
      # This is a dataframe - process it
      x %>% 
        {if ("trial_n" %in% names(.)) select(., -trial_n) else .} %>%
        mutate(trials = 1:n())
    } else if (is.list(x)) {
      # This is a list - recurse deeper
      lapply(x, modify_structure)
    } else {
      # This is something else (character, etc.) - leave unchanged
      x
    }
  }
  
  modified_data <- modify_structure(split_data)
  return(modified_data)
}

split_results <- process_split_trials_safe(split_results)

```

## Merging each of these into a big data set 


```{r}

#extract all data frames
extract_dfs <- function(x) {
  if (is.data.frame(x)) {
    list(x)
  } else if (is.list(x)) {
    unlist(lapply(x, extract_dfs), recursive = FALSE)
  } else {
    NULL
  }
}

#Un-nest the list
all_dfs <- extract_dfs(split_results)

#Everythings become one
Big_data <- do.call(rbind, lapply(seq_along(all_dfs), function(i) {
  cbind(all_dfs[[i]], group_id = names(all_dfs)[i])
}))


```

# Now we can extract what interest us the most which is the final Threshold, Contrast for each durations

```{r}
Big_data <- subset(na.omit(Big_data))
Big_data <- subset(Big_data, updated_threshold < 99)
Final_data <- subset(Big_data, trials == max(trials))

#We can aslo add a Energy column which is the product of Duration and contrast

Final_data$duration <- as.numeric(as.character(Final_data$duration))
energy <- c(Final_data$duration * Final_data$contrast)
Final_data <- data.frame(Final_data, energy)


#plot(Final_data$updated_threshold~Final_data$duration)



den_plot <- ggplot(data= Final_data, aes(x=updated_threshold)) + geom_density()

den_plot

ploploplot <- ggplot(data = Final_data, aes(y = energy, x =duration)) + geom_jitter(aes(color = task)) + theme_classic()
ploploplot

logploplot <- ggplot(data = Final_data, aes(y = log(energy), x = log(duration))) + geom_jitter(aes(color = task)) + theme_classic()
logploplot

```
I still need to add some plot similar to the one in the contrast task script to check if any participants is doing anything out the ordinary.
