---
title: "PSE_Bloch_law_plot"
author: "oscar"
date: "2025-06-17"
output: html_document
---

# Import the Data and organize them
First we import the data of each tasks for one participant and each r library that we need.
```{r, echo=TRUE}
library(readr)
library(ggplot2)
library(dplyr)
library(ggrepel)

#Change the number xx_faces_discr/BL_faces_xx_data_table.csv" and re-run all chunck to see the plot of another participant number xx


contr_landscape <- read_csv("/home/ruiz/Documents/Stage_brxl/DATA_BOCH_LAW/PSE/subj_01_bloch_law_landscapes_PSE/BL_landscapes_PSE_01_data_table.csv")
contr_faces <- read_csv("/home/ruiz/Documents/Stage_brxl/DATA_BOCH_LAW/PSE/subj_01_bloch_law_faces_PSE/BL_faces_PSE_01_data_table.csv")
contr_objects <- read_csv("/home/ruiz/Documents/Stage_brxl/DATA_BOCH_LAW/PSE/subj_01_bloch_law_objects_PSE/BL_objects_PSE_01_data_table.csv")

# we use head to check if each data set are imported properly.
head(contr_objects)
head(contr_faces)
head(contr_landscape)
```

## Dividing the data frames into subsets

We divide the three data frames into subsets for eachs 10 durations and we extract the final threshold for each of theses durations and we also extract the final contrast for each of theses durations.
```{r, echo=FALSE}
split_by_factor <- function(data, factor_var) {
  # Convert variable to factor if it isn't already
  if (!is.factor(data[[factor_var]])) {
    data[[factor_var]] <- factor(data[[factor_var]])
  }
  
  # Split data frame into list of subsets
  split(data, data[[factor_var]])
}


subsets_faces <- split_by_factor(contr_faces, "duration_test")
subsets_objects <- split_by_factor(contr_objects, "duration_test")
subsets_landscapes <- split_by_factor(contr_landscape, "duration_test")

# I tried to create a function to replace each column trial_n of each subsets by the vector "trials" but because every columns dont have the same length it create an error that I'm too lazy  to fix for now. (it still work)
trials <- seq(from = 1, to = 50 , length = 50)
for (i in 1:10) {
  subsets_faces[[i]]$trial_n <- trials
}
for (i in 1:10) {
  subsets_objects[[i]]$trial_n <- trials
}
for (i in 1:10) {
  subsets_landscapes[[i]]$trial_n <- trials
}

#Now we create a vector containing the last updated value for each duration and a vector of the contrast for each of these value

final_thresholds_faces <- sapply(1:10, function(i) subsets_faces[[i]]$updated_threshold[50])
final_thresholds_objects <- sapply(1:10, function(i) subsets_objects[[i]]$updated_threshold[50])
final_thresholds_landscapes <- sapply(1:10, function(i) subsets_landscapes[[i]]$updated_threshold[50])


final_contrasts_faces <- sapply(1:10, function(i) subsets_faces[[i]]$contrast_test[50])
final_contrasts_objects <- sapply(1:10, function(i) subsets_objects[[i]]$contrast_test[50])
final_contrasts_landscapes <- sapply(1:10, function(i) subsets_landscapes[[i]]$contrast_test[50])
```

# Stairs plots
Now we create plot of updated threshold for each trials of each duration to see if the algorithm is working. If it worked we should see stair like plot that converge to a certain value at the 50th trials. If we dont see any stair shape it mean that it didn't work properly, and if the value to which the updated threshold converge is 100 or > 99 it mean that we can't find the "true" threshold wihin our contrasts range. So theses value will be filtered later.

```{r, echo=FALSE,fig.show="hold", out.width="50%"}
plot_stair_faces <- function(data, duration_level, threshold_50){
  ggplot(data = data , aes(x = trial_n)) +
    geom_point(aes(y = updated_threshold, colour = factor(resp))) +
    geom_line(aes(y = updated_threshold, linetype = "Threshold"), color = "black") +
    geom_line(aes(y = contrast_test, linetype = "Contrast"), color = "black") +
    scale_linetype_manual(name = "Lines",
                        values = c("Threshold" = "solid", "Contrast" = "dashed")) +
    scale_color_manual(name = "Response",
                       values = c("0" = "red", "1" = "lightgreen"),  # Adjust these colors as needed
                       labels = c("0" = "Incorrect", "1" = "Correct")) +
    theme_minimal() +
    labs(title = paste("Plot of the predicted threshold for each trial for a duration (µs) of ",levels(factor(duration_level))[i]),
         subtitle = "Faces task",
         x = "Trials number",
         y = "Predicted threshold/Contrast") +
    theme(legend.position = "right",
          legend.box = "vertical")+
    geom_hline(yintercept = threshold_50, linetype = "dashed", color = "red")
}

plot_stair_objects <- function(data, duration_level, threshold_50){
  ggplot(data = data , aes(x = trial_n)) +
    geom_point(aes(y = updated_threshold, colour = factor(resp))) +
    geom_line(aes(y = updated_threshold, linetype = "Threshold"), color = "black") +
    geom_line(aes(y = contrast_test, linetype = "Contrast"), color = "black") +
    scale_linetype_manual(name = "Lines",
                        values = c("Threshold" = "solid", "Contrast" = "dashed")) +
    scale_color_manual(name = "Response",
                       values = c("0" = "red", "1" = "lightgreen"),  # Adjust these colors as needed
                       labels = c("0" = "Incorrect", "1" = "Correct")) +
    theme_minimal() +
    labs(title = paste("Plot of the predicted threshold for each trial for a duration (µs) of ",levels(factor(duration_level))[i]),
         subtitle = "Objects task",
         x = "Trials number",
         y = "Predicted threshold/Contrast") +
    theme(legend.position = "right",
          legend.box = "vertical")+
    geom_hline(yintercept = threshold_50, linetype = "dashed", color = "red")
}

plot_stair_landscapes <- function(data, duration_level, threshold_50){
  ggplot(data = data , aes(x = trial_n)) +
    geom_point(aes(y = updated_threshold, colour = factor(resp))) +
    geom_line(aes(y = updated_threshold, linetype = "Threshold"), color = "black") +
    geom_line(aes(y = contrast_test, linetype = "Contrast"), color = "black") +
    scale_linetype_manual(name = "Lines",
                        values = c("Threshold" = "solid", "Contrast" = "dashed")) +
    scale_color_manual(name = "Response",
                       values = c("0" = "red", "1" = "lightgreen"),  # Adjust these colors as needed
                       labels = c("0" = "Incorrect", "1" = "Correct")) +
    theme_minimal() +
    labs(title = paste("Plot of the predicted threshold for each trial for a duration (µs) of ",levels(factor(duration_level))[i]),
         subtitle = "Landscapes task",
         x = "Trials number",
         y = "Predicted threshold/Contrast") +
    theme(legend.position = "right",
          legend.box = "vertical")+
    geom_hline(yintercept = threshold_50, linetype = "dashed", color = "red")
}

par(mar = c(6, 5, .1, .1))
#We then loop each of the three function above to create 10 plots for each task (one plot per durations per task)
for (i in 1:10) {
  print(
    plot_stair_faces(
      data = subsets_faces[[i]],
      duration_level = levels(factor(contr_faces$duration_test))[i],
      threshold_50 = final_thresholds_faces[i]
    )
  )
}

for (i in 1:10) {
  print(
    plot_stair_objects(
      data = subsets_objects[[i]],
      duration_level = levels(factor(contr_objects$duration_test))[i],
      threshold_50 = final_thresholds_objects[i]
    )
  )
}

for (i in 1:10) {
  print(
    plot_stair_landscapes(
      data = subsets_landscapes[[i]],
      duration_level = levels(factor(contr_landscape$duration_test))[i],
      threshold_50 = final_thresholds_landscapes[i]
    )
  )
}
```


```{r}


final_thresholds_faces <- sapply(1:10, function(i) subsets_faces[[i]]$updated_threshold[50])
final_thresholds_objects <- sapply(1:10, function(i) subsets_objects[[i]]$updated_threshold[50])
final_thresholds_landscapes <- sapply(1:10, function(i) subsets_landscapes[[i]]$updated_threshold[50])


final_contrasts_faces <- sapply(1:10, function(i) subsets_faces[[i]]$contrast_test[50])
final_contrasts_objects <- sapply(1:10, function(i) subsets_objects[[i]]$contrast_test[50])
final_contrasts_landscapes <- sapply(1:10, function(i) subsets_landscapes[[i]]$contrast_test[50])
DurationF <- c(500, 901, 1623, 2924, 5268, 9491, 17100, 30808, 55505, 1e+05)


data_faces <- data.frame(final_thresholds_faces,final_contrasts_faces,DurationF)
Fil_data_faces <- subset(data_faces, final_thresholds_faces < 99)

data_objects <- data.frame(final_thresholds_objects,final_contrasts_objects,DurationF)
Fil_data_objects <- subset(data_objects, final_thresholds_faces < 99)

data_landscapes <- data.frame(final_thresholds_landscapes,final_contrasts_landscapes,DurationF)
Fil_data_landscapes <- subset(data_landscapes, final_thresholds_landscapes < 99)

inverse_relationship_faces <- ggplot(data = Fil_data_faces, aes(final_contrasts_faces, final_thresholds_faces,label = factor(DurationF)))  + geom_point() + geom_text_repel() + theme_classic() + 
  labs(title = "final Threshold depending of the final contrast", 
       subtitle = "the numbers indicate the duration in µs" ,
       x = "Contrast", 
       y = "Final treshold",
       caption = "faces task")

inverse_relationship_objects<- ggplot(data = Fil_data_objects, aes(final_contrasts_objects, final_thresholds_objects,label = factor(DurationF)))  + geom_point() + geom_text_repel() + theme_classic() + 
  labs(title = "final Threshold depending of the final contrast",
       subtitle = "the numbers indicate the duration in µs" , 
       x = "Contrast", 
       y = "Final treshold", 
       caption = "objects task")

inverse_relationship_landscapes <- ggplot(data = Fil_data_landscapes, aes(final_contrasts_landscapes, final_thresholds_landscapes,label = factor(DurationF)))  + geom_point() + geom_text_repel() + theme_classic() + 
  labs(title = "final Threshold depending of the final contrast",
       subtitle = "the numbers indicate the duration in µs" ,
       x = "Contrast",
       y = "Final treshold", 
       caption = "landscapes task")





inverse_relationship_faces
inverse_relationship_landscapes
inverse_relationship_objects

```

```{r}
duration_contrast_l <- ggplot(data = Fil_data_landscapes, aes(final_contrasts_landscapes, DurationF))  + geom_point()  + theme_classic() + 
  labs(title = "final Threshold~duration",
       x = "Duration",
       y = "Final treshold", 
       caption = "landscapes task")

duration_contrast_f <- ggplot(data = Fil_data_faces, aes(final_contrasts_faces, DurationF))  + geom_point()  + theme_classic() + 
  labs(title = "final Threshold~duration",
       x = "Duration",
       y = "Final treshold", 
       caption = "landscapes task")

duration_contrast_l
duration_contrast_f
head(Fil_data_faces)
```

```{r}

Energy_duration_plot_l <- ggplot(Fil_data_landscapes, aes(x = DurationF, y = final_contrasts_landscapes*DurationF)) + geom_point() + labs(title = "Plot of Energy (duration x contrast) by duration ", y = "Energy", x="Durations", subtitle = "Landscapes task") +theme_classic()

LogEnergy_Logduration_plot_l <- ggplot(Fil_data_landscapes, aes(x = log(DurationF), y = log(final_contrasts_landscapes*DurationF))) + geom_point() + labs(title = "Plot of Energy (duration x contrast) by duration ", y = "log(Energy)", x="log(Duration)", subtitle = "Landscapes task") +theme_classic()

Energy_duration_plot_l
LogEnergy_Logduration_plot_l

Energy_duration_plot_f <- ggplot(Fil_data_faces, aes(x = DurationF, y = final_contrasts_faces*DurationF)) + geom_point() + labs(title = "Plot of Energy (duration x contrast) by duration ", y = "Energy", x="Durations", subtitle = "Faces task") +theme_classic()

LogEnergy_Logduration_plot_f <- ggplot(Fil_data_faces, aes(x = log(DurationF), y = log(final_contrasts_faces*DurationF))) + geom_point() + labs(title = "Plot of Energy (duration x contrast) by duration ", y = "log(Energy)", x="log(Duration)", subtitle = "Faces task") +theme_classic()

Energy_duration_plot_f
LogEnergy_Logduration_plot_f

Energy_duration_plot_o <- ggplot(Fil_data_objects, aes(x = DurationF, y = final_contrasts_objects*DurationF)) + geom_point() + labs(title = "Plot of Energy (duration x contrast) by duration ", y = "Energy", x="Durations", subtitle = "objects task") +theme_classic()

LogEnergy_Logduration_plot_o <- ggplot(Fil_data_objects, aes(x = log(DurationF), y = log(final_contrasts_objects*DurationF))) + geom_point() + labs(title = "Plot of Energy (duration x contrast) per duration ", y = "log(Energy)", x="log(Duration)", subtitle = "objects task") +theme_classic()

Energy_duration_plot_o
LogEnergy_Logduration_plot_o
```